---
title: "Plots of model fit for all 16 models, for all 3 pgroups"
output:
  html_document: default
  pdf_document: default
date: "2025-05-23"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Source

Takes model and data in csv `fitforplot16m.csv` calculated in `optimise_withKandEps.Rmd`. 


## Note on parameters

Predictions generated from the best fitting parameter combination for each model.

All models have epsilon and tau. First eight have kappa (to go with Kindness).

```{r, include=FALSE}
library(RColorBrewer)
library(tidyverse)
library(rlang)
library(purrr)
```

```{r, echo=FALSE}
df <- read.csv('../model_data/fitforplot16m.csv') # 288 obs of 27 vars - only the actual values of A/B
```


```{r, include=FALSE}
df$structure <- ifelse(substr(df$trial_id, 3, 3) == "c", "conjunctive", 
                    ifelse(substr(df$trial_id, 3, 3) == "d", "disjunctive", NA))


df$trial_type <- recode(df$trial_id, 
                      "1_c1"="A=0,B=0,E=0",
                      "1_c2"="A=0,B=1,E=0",
                      "1_c3"="A=1,B=0,E=0",
                      "1_c4"="A=1,B=1,E=0",
                      "1_c5"="A=1,B=1,E=1",
                      "1_d1"="A=0,B=0,E=0",
                      "1_d2"="A=0,B=1,E=0",
                      "1_d3"="A=0,B=1,E=1",
                      "1_d4"="A=1,B=0,E=0",
                      "1_d5"="A=1,B=0,E=1",
                      "1_d6"="A=1,B=1,E=0",
                      "1_d7"="A=1,B=1,E=1",
                      "2_c1"="A=0,B=0,E=0",
                      "2_c2"="A=0,B=1,E=0",
                      "2_c3"="A=1,B=0,E=0",
                      "2_c4"="A=1,B=1,E=0",
                      "2_c5"="A=1,B=1,E=1",
                      "2_d1"="A=0,B=0,E=0",
                      "2_d2"="A=0,B=1,E=0",
                      "2_d3"="A=0,B=1,E=1",
                      "2_d4"="A=1,B=0,E=0",
                      "2_d5"="A=1,B=0,E=1",
                      "2_d6"="A=1,B=1,E=0",
                      "2_d7"="A=1,B=1,E=1",
                      "3_c1"="A=0,B=0,E=0",
                      "3_c2"="A=0,B=1,E=0",
                      "3_c3"="A=1,B=0,E=0",
                      "3_c4"="A=1,B=1,E=0",
                      "3_c5"="A=1,B=1,E=1",
                      "3_d1"="A=0,B=0,E=0",
                      "3_d2"="A=0,B=1,E=0",
                      "3_d3"="A=0,B=1,E=1",
                      "3_d4"="A=1,B=0,E=0",
                      "3_d5"="A=1,B=0,E=1",
                      "3_d6"="A=1,B=1,E=0",
                      "3_d7"="A=1,B=1,E=1")

df$trial_structure_type <- recode(df$trial_id, 
                                "1_c1"="Conjunctive: A=0,B=0,E=0",
                                "1_c2"="Conjunctive: A=0,B=1,E=0",
                                "1_c3"="Conjunctive: A=1,B=0,E=0",
                                "1_c4"="Conjunctive: A=1,B=1,E=0",
                                "1_c5"="Conjunctive: A=1,B=1,E=1",
                                "1_d1"="Disjunctive: A=0,B=0,E=0",
                                "1_d2"="Disjunctive: A=0,B=1,E=0",
                                "1_d3"="Disjunctive: A=0,B=1,E=1",
                                "1_d4"="Disjunctive: A=1,B=0,E=0",
                                "1_d5"="Disjunctive: A=1,B=0,E=1",
                                "1_d6"="Disjunctive: A=1,B=1,E=0",
                                "1_d7"="Disjunctive: A=1,B=1,E=1",
                                "2_c1"="Conjunctive: A=0,B=0,E=0",
                                "2_c2"="Conjunctive: A=0,B=1,E=0",
                                "2_c3"="Conjunctive: A=1,B=0,E=0",
                                "2_c4"="Conjunctive: A=1,B=1,E=0",
                                "2_c5"="Conjunctive: A=1,B=1,E=1",
                                "2_d1"="Disjunctive: A=0,B=0,E=0",
                                "2_d2"="Disjunctive: A=0,B=1,E=0",
                                "2_d3"="Disjunctive: A=0,B=1,E=1",
                                "2_d4"="Disjunctive: A=1,B=0,E=0",
                                "2_d5"="Disjunctive: A=1,B=0,E=1",
                                "2_d6"="Disjunctive: A=1,B=1,E=0",
                                "2_d7"="Disjunctive: A=1,B=1,E=1",
                                "3_c1"="Conjunctive: A=0,B=0,E=0",
                                "3_c2"="Conjunctive: A=0,B=1,E=0",
                                "3_c3"="Conjunctive: A=1,B=0,E=0",
                                "3_c4"="Conjunctive: A=1,B=1,E=0",
                                "3_c5"="Conjunctive: A=1,B=1,E=1",
                                "3_d1"="Disjunctive: A=0,B=0,E=0",
                                "3_d2"="Disjunctive: A=0,B=1,E=0",
                                "3_d3"="Disjunctive: A=0,B=1,E=1",
                                "3_d4"="Disjunctive: A=1,B=0,E=0",
                                "3_d5"="Disjunctive: A=1,B=0,E=1",
                                "3_d6"="Disjunctive: A=1,B=1,E=0",
                                "3_d7"="Disjunctive: A=1,B=1,E=1")

df<- df %>% mutate(trial_type = factor(trial_type, levels = c("A=0,B=0,E=0",
                                                              "A=0,B=1,E=0",
                                                              "A=0,B=1,E=1",
                                                              "A=1,B=0,E=0",
                                                              "A=1,B=0,E=1",
                                                              "A=1,B=1,E=0",
                                                              "A=1,B=1,E=1")),
                   trial_structure_type = factor(trial_structure_type, levels = c("Conjunctive: A=1,B=1,E=1",
                                                                                  "Conjunctive: A=1,B=1,E=0",
                                                                                  "Conjunctive: A=1,B=0,E=0",
                                                                                  "Conjunctive: A=0,B=1,E=0",
                                                                                  "Conjunctive: A=0,B=0,E=0",
                                                                                  "Disjunctive: A=0,B=1,E=0",
                                                                                  "Disjunctive: A=1,B=1,E=1",
                                                                                  "Disjunctive: A=1,B=0,E=1",
                                                                                  "Disjunctive: A=0,B=1,E=1",
                                                                                  "Disjunctive: A=1,B=1,E=0",
                                                                                  "Disjunctive: A=1,B=0,E=0",
                                                                                  "Disjunctive: A=0,B=0,E=0")),
                   # NOTE THESE (ABOVE) LOOK IN THE WRONG ORDER BUT ITS JUST FOR THE PLOT
                   pgroup = factor(pgroup, levels = 1:3, labels = c('A=.1,Au=.5,B=.8,Bu=.5',
                                                                            'A=.5,Au=.1,B=.5,Bu=.8',
                                                                            'A=.1,Au=.7,B=.8,Bu=.5')),
                   Response = factor(node3),
                   Observed = factor(!Response%in%c('Au=0','Au=1','Bu=0', 'Bu=1'), levels = c(T,F)),
                   Variable = factor(substr(node3, 1, 1)),
                   #State = factor(rep(c(0,1), 144)),
                   Actual = factor(Actual, levels = c(FALSE, TRUE), labels = c('FALSE', 'TRUE')))

```



```{r, echo=FALSE}
df$SE <- NA

# Need to add SE 
for (i in unique(df$trial_id))
{
  df$SE[df$trial_id==i]<-sqrt((df$prop[df$trial_id==i] * (1-df$prop[df$trial_id==i])) / sum(df$n[df$trial_id==i]))
}

```


## Lesions and order

Models have four 'components': *Kindness*, *Selection*, *Inference*, *Actual causation*.

The full model has all four.

Thereafter they are progressively lesioned.

They are lesioned in the following way:

- First block of eight inc. full: has Kindness, with all other modules lesioned dissociatively (naming convention `noX' means lesioned to remove the X module.

- Next block of eight: follows the same process in order, this time without Kindness.


```{r, echo=FALSE}

models <- c('full', 
                 'noAct', 
                 'noInf', 
                 'noSelect', 
                 'noActnoInf', 
                 'noActnoSelect', 
                 'noInfnoSelect', 
                 'noActnoInfnoSelect',
                 'noKind', 
                 'noActnoKind', 
                 'noInfnoKind', 
                 'noKindnoSelect', 
                 'noActnoInfnoKind', 
                 'noActnoKindnoSelect', 
                 'noInfnoKindnoSelect', 
                 'noActnoInfnoKindnoSelect')  

pgroups <- levels(df$pgroup)


```



```{r, include=FALSE}

# Define plotting function
plot_model_pgroup <- function(model_colname, pgroup_label, df) {
df_filtered <- df %>% filter(pgroup == pgroup_label)
ggplot(df_filtered, aes(x = node3, y = prop, fill = Observed, colour = Actual)) +
geom_bar(stat = 'identity') +
geom_errorbar(aes(ymin = prop-SE, ymax = prop+SE), width = .2) +
labs(
x = 'Response',
y = 'Proportion/Prediction',
title = sprintf("%s : %s", model_colname, pgroup_label)
) +
scale_fill_brewer(palette = "Set2") +
scale_colour_manual(values = c('gray', 'black')) +
geom_point(aes(y = .data[[model_colname]]), colour = 'black') +
geom_rect(
data = subset(df_filtered, trial_structure_type %in% c("Conjunctive: A=1,B=1,E=1", "Disjunctive: A=1,B=1,E=1")),
fill = NA, colour = "blue", xmin = -Inf, xmax = Inf, linewidth = 2,
ymin = -Inf, ymax = Inf
) +
facet_wrap(~trial_structure_type, ncol = 6) +
theme_bw() +
theme(
panel.grid = element_blank(),
axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1),
legend.margin = margin(c(0, 0, 0, 0)),
axis.title.x = element_text(margin = margin(t = 1, r = 0, b = 0, l = 0)),
plot.title = element_text(size = 18, face = "bold", hjust = 0.5)
)
}
```


```{r, echo=FALSE, fig.width=12}
for (model in models) {
for (pgroup in pgroups) {
print(plot_model_pgroup(model, pgroup, df))
}
}
```


A different function just for one plot:

```{r, echo=FALSE, fig.width=12}
plot_two_models_pgroup <- function(model1, model2, pgroup_label, df) {
  df_filtered <- df %>% filter(pgroup == pgroup_label)
  ggplot(df_filtered, aes(x = node3, y = prop, fill = Observed, colour = Actual)) +
    geom_bar(stat = 'identity') +
    geom_errorbar(aes(ymin = prop-SE, ymax = prop+SE), width = .2) +
    # Model 1: black circles
    geom_point(aes(y = .data[[model1]]), 
               colour = 'black', 
               shape = 16, 
               size = 2, 
               alpha = 0.9,
               show.legend = TRUE) +
    # Model 2: red triangles
    geom_point(aes(y = .data[[model2]]), 
               colour = 'red', 
               shape = 17, 
               size = 2, 
               alpha = 0.9,
               show.legend = TRUE) +
    geom_rect(
      data = subset(df_filtered, trial_structure_type %in% c("Conjunctive: A=1,B=1,E=1", "Disjunctive: A=1,B=1,E=1")),
      fill = NA, colour = "blue", xmin = -Inf, xmax = Inf, linewidth = 2,
      ymin = -Inf, ymax = Inf
    ) +
    facet_wrap(~trial_structure_type, ncol = 6) +
    labs(
      x = 'Response',
      y = 'Proportion/Prediction',
      title = sprintf("%s (black) & %s (red) - Parameter Group: %s", model1, model2, pgroup_label)
    ) +
    scale_fill_brewer(palette = "Set2") +
    scale_colour_manual(values = c('gray', 'black')) +
    theme_bw() +
    theme(
      panel.grid = element_blank(),
      axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1),
      legend.margin = margin(c(0, 0, 0, 0)),
      axis.title.x = element_text(margin = margin(t = 1, r = 0, b = 0, l = 0))
    )
}
```


Usage:

```{r, echo=FALSE, fig.width=12}
# For example, to compare 'full' and 'noSelect' for a given pgroup:
p <- plot_two_models_pgroup('full', 'noSelect', 'A=.1,Au=.7,B=.8,Bu=.5', df)

ggsave("model_comparison.pdf", plot = p, width = 7, height = 5, units = "in")

```


Now just the 111s

```{r}
ggplot(df %>% filter(trial_structure_type %in% c("Conjunctive: A=1,B=1,E=1", "Disjunctive: A=1,B=1,E=1")),
       aes(x=node3, y=prop, fill=Observed, colour = Actual)) +
         geom_bar(stat = 'identity') +
         geom_errorbar(aes(ymin=prop-SE, ymax=prop+SE), width=.2) + 
         labs(x = 'Response', y = 'Proportion/Prediction')+
         scale_fill_brewer(palette = "Set2")+
         scale_colour_manual(values = c('black', 'gray')) +
         geom_point(aes(y=full), colour = 'black') +
         # geom_point(aes(y=noInf), colour = 'black', shape = 2) +
         # geom_point(aes(y=noSelect), colour = 'black', shape = 3) +
         facet_grid(trial_structure_type ~ prob_group) +
         theme_bw() +
         theme(panel.grid = element_blank(),
               axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1),
               legend.margin=margin(c(0,0,0,0)),
               axis.title.x = element_text(margin = margin(t = 1, r =0, b = 0, l = 0))) 
ggsave('results_111.pdf', width = 7, height = 5)

```

Looking for the unobserved only!

```{r}
# 48 obs of 4 - that is: 2(obs/un) * 2(ppts/full) * 12 trialtypes
df.l <- df %>% gather(key, val, prop:full) %>% mutate(key = factor(key, levels = c('prop', 'full'),
                                                                 labels = c("Participants", "Full Model"))) %>%
  group_by(Observed, key, trial_structure_type) %>% summarise(val = sum(val)/3)

df.l <- df.l %>% filter(!is.na(key))
```

A plot for the FULL MODEL to compare unobserved with observed variables:

```{r}
# Now same thing for structure only
df.s <- df %>% gather(key, val, prop:full) %>% mutate(key = factor(key, levels = c('prop', 'full'),
                                                                 labels = c("Participants", "Full Model"))) %>%
  group_by(Observed, key, structure, E) %>% summarise(val = sum(val)/3)

df.s <- df.s %>% filter(!is.na(key))

df.s$struct_effect <- paste0(df.s$structure, df.s$E)

# This is for with no Effect, use a different line for when include effect
#df.s$val <- ifelse(df.s$structure == "conjunctive", df.s$val/5,
                    #ifelse(df.s$structure == "disjunctive", df.s$val/7, NA))

# Better
df.s$val <- ifelse(df.s$struct_effect == "conjunctive0", df.s$val/4,
                    ifelse(df.s$struct_effect == "disjunctive0", df.s$val/4,
                           ifelse(df.s$struct_effect == "disjunctive1", df.s$val/3, df.s$val)))

```

This is the two-panel plot:

```{r}
ggplot(df.s, aes(y=val, x=key, fill=Observed)) +
  stat_summary(fun.y = mean,
               geom = "bar",
               position = position_dodge(), colour = 'black') +
stat_summary(fun.data = mean_se, geom = "errorbar",  position = position_dodge(.9), width = .2)+
  labs(x = 'Response', y = 'Proportion/Prediction') +
  facet_grid(structure ~ E) +
  scale_fill_brewer(palette = "Set2") +
  theme_bw() +
  theme(panel.grid = element_blank(),
        axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1),
        legend.margin=margin(c(0,0,0,0)),
        #legend.position = c(-.1,-.2),
        axis.title.x = element_text(margin = margin(t = 1, r =0, b = 0, l = 0))) 

```

Now the same thing but for different models:


```{r}
# Now same thing for structure only
df.s2 <- df %>% gather(key, val, prop:noKind) %>% mutate(key = factor(key, levels = c('prop', 'noKind'),
                                                                 labels = c("Participants", "noKind Model"))) %>%
  group_by(Observed, key, structure, E) %>% summarise(val = sum(val)/3)

df.s2 <- df.s2 %>% filter(!is.na(key))

df.s2$struct_effect <- paste0(df.s2$structure, df.s2$E)

# This is for with no Effect, use a different line for when include effect
#df.s$val <- ifelse(df.s$structure == "conjunctive", df.s$val/5,
                    #ifelse(df.s$structure == "disjunctive", df.s$val/7, NA))

# Better
df.s2$val <- ifelse(df.s2$struct_effect == "conjunctive0", df.s2$val/4,
                    ifelse(df.s2$struct_effect == "disjunctive0", df.s2$val/4,
                           ifelse(df.s2$struct_effect == "disjunctive1", df.s2$val/3, df.s2$val)))

```

This is the two-panel plot:

```{r}
ggplot(df.s2, aes(y=val, x=key, fill=Observed)) +
  stat_summary(fun.y = mean,
               geom = "bar",
               position = position_dodge(), colour = 'black') +
stat_summary(fun.data = mean_se, geom = "errorbar",  position = position_dodge(.9), width = .2)+
  labs(x = 'Response', y = 'Proportion/Prediction', title = 'noKind model') +
  facet_grid(structure ~ E) +
  scale_fill_brewer(palette = "Set2") +
  theme_bw() +
  theme(panel.grid = element_blank(),
        axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1),
        legend.margin=margin(c(0,0,0,0)),
        #legend.position = c(-.1,-.2),
        axis.title.x = element_text(margin = margin(t = 1, r =0, b = 0, l = 0))) 

```





Rename and rerun this bit once you're happy with the plots:

```{r}
ggsave('results_obs_un.pdf', width = 7, height = 5)    
```

Now also do this for noSelect, and the full model without Kindness.

Also split out E=0 and E=1. For this, need to make a new variable





```{r}
row_labeller <- c(
  "A=.1,Au=.5,B=.8,Bu=.5"=".1,.5,.8,.5",
  "A=.5,Au=.1,B=.5,Bu=.8"=".5,.1,.5,.8",
  "A=.1,Au=.7,B=.8,Bu=.5"=".1,.7,.8,.5"
)

column_labeller <- c(
  "Conjunctive: A=1,B=1,E=1"="Conjunctive",
  "Disjunctive: A=1,B=1,E=1"="Disjunctive"
)


ggplot(df %>% filter(trial_structure_type %in% c("Conjunctive: A=1,B=1,E=1", "Disjunctive: A=1,B=1,E=1"), Actual == T),
       aes(x=Variable, y=prop, fill=Observed)) +
  geom_bar(stat = 'identity', colour = 'black', position = position_dodge()) +
  geom_errorbar(aes(ymin=prop-SE, ymax=prop+SE), width=.2, position = position_dodge(.9)) + 
  labs(x = 'Response', y = 'Proportion/Prediction')+
  scale_fill_brewer(palette = "Set2")+
  geom_point(data = df %>% filter(trial_structure_type %in% c("Conjunctive: A=1,B=1,E=1", "Disjunctive: A=1,B=1,E=1"), Actual==T), aes(y=full), colour = 'black', position = position_dodge(.9)) +
  # geom_point(aes(y=noInf), colour = 'black', shape = 2) +
  # geom_point(aes(y=noSelect), colour = 'black', shape = 3) +
  facet_grid(trial_structure_type ~ pgroup,
             labeller = labeller(pgroup = column_labeller, trial_structure_type = row_labeller)) +
  theme_bw() +
  theme(panel.grid = element_blank(),
        legend.position = 'none',
        axis.title.x = element_text(margin = margin(t = 1, r =0, b = 0, l = 0))) 
ggsave('results_simple_111.pdf', width = 7, height = 5)        

```